---
title: "DATA 555 Final Project"
author: "Annie Nicholas"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
```

Column {.sidebar data-width=300}
-------------------------------------

### **Study Information**

Individual participant data meta-regression was used to evaluate the reactogenicity-immunogenicity relationship among 12 pandemic influenza vaccines. Data comes from clinical trials conducted by the National Institute of Allergy and Infectious Diseases and is not publicly available. Trials took place between 2009 and 2020 at Vaccine and Treatment Evaluation Unit network locations. Participants were healthy, non-pregnant, immuno-competent adults. 

Note: All 12 studies were used for bivariable analysis, but only 11 studies were included in the full model.

**Covariates included in the model:**

- Sex 
- Age group (18-30, 31-40, 41-50, 51-64, 65+ years old)
- Vaccine dosage (low, medium, high)
- Adjuvant (combinations of AS03 and MF59)
- Strain (H5N1, H5N8, H7N9)

Tip: In the forest plot, click on a strain in the legend to hide it on the plot.

Row {.tabset data-width=700 data-height=800}
-------------------------------------

### Bivariable Association

```{r}
library(ggplot2)
library(dplyr)
library(lme4)
library(ggpubr)
library(janitor)
library(RColorBrewer)
library(plotly)
library(ggpp)

d <- read.csv("data/Final_1113.csv")
d <- d[!is.na(d$final_logratio), ]

d <- d %>%
  mutate(dosage = case_when(
    total_dose <= 11.25 ~ "low",
    total_dose <= 22.5 ~ "med",
    TRUE ~ "high"
  ))

d <- d %>%
  mutate(adj = case_when(
    study == "09-0058" & trttrue == "T1" ~ 1,
    study == "09-0058" & trttrue == "T2" ~ 1,
    study == "09-0058" & trttrue == "T3" ~ 1,
    study == "10-0017" & trttrue == "T1" ~ 1,
    study == "10-0017" & trttrue == "T2" ~ 1,
    study == "10-0017" & trttrue == "T3" ~ 1,
    study == "13-0033" & trttrue == "T1" ~ 1,
    study == "13-0033" & trttrue == "T2" ~ 1,
    study == "13-0033" & trttrue == "T3" ~ 1,
    study == "15-0064" & trttrue == "1" ~ 1,
    study == "15-0064" & trttrue == "4" ~ 1,
    study == "15-0066" & trttrue == "GROUP1" ~ 1,
    study == "15-0066" & trttrue == "GROUP2" ~ 1,
    study == "17-0075" & trttrue == "GROUP1" ~ 1,
    study == "17-0075" & trttrue == "GROUP2" ~ 1,
    study == "17-0075" & trttrue == "GROUP3" ~ 1,
    study == "17-0078" & trttrue == "1" ~ 1,
    study == "17-0078" & trttrue == "2" ~ 1,
    study == "17-0078" & trttrue == "4" ~ 1,
    study == "17-0078" & trttrue == "5" ~ 1,
    study == "17-0090" & trttrue == "1" ~ 1,
    study == "10-0016" & trttrue == "T1" ~ 2,
    study == "10-0016" & trttrue == "T2" ~ 2,
    study == "10-0016" & trttrue == "T3" ~ 2,
    study == "13-0032" & trttrue == "T1" ~ 2,
    study == "13-0032" & trttrue == "T2" ~ 2,
    study == "13-0032" & trttrue == "T3" ~ 2,
    study == "13-0033" & trttrue == "T8" ~ 2,
    study == "13-0034" ~ 2,
    study == "15-0064" & trttrue == "2" ~ 2,
    study == "15-0064" & trttrue == "5" ~ 2,
    study == "15-0066" & trttrue == "GROUP4" ~ 2,
    study == "15-0066" & trttrue == "GROUP5" ~ 2,
    study == "17-0076" & trttrue == "3.7579WA2x" ~ 2,
    study == "17-0076" & trttrue == "7.579WA2x" ~ 2,
    study == "17-0076" & trttrue == "1579WA2x" ~ 2,
    study == "13-0033" & trttrue == "T4" ~ 3,
    study == "17-0078" & trttrue == "3" ~ 3,
    study == "17-0078" & trttrue == "6" ~ 3,
    study == "13-0033" & trttrue == "T5" ~ 3,
    study == "13-0032" & trttrue == "T4" ~ 4,
    study == "13-0032" & trttrue == "T5" ~ 4,
    study == "13-0033" & trttrue == "T6" ~ 5,
    study == "13-0033" & trttrue == "T7" ~ 5,
    TRUE ~ 0
  ))

# More detailed adj variable
# 0 = no adj
# 1 = none then AS03
# 2 = none then MF59
# 3 = AS03 then none
# 4 = MF59 then none
# 5 = AS03 then MF59
# 6 = MF59 then AS03
# 7 = AS03 for all
# 8 = MF59 for all

d <- d %>%
  mutate(adj_2 = case_when(
    study == "09-0058" & trttrue == "T1" ~ 7,
    study == "09-0058" & trttrue == "T2" ~ 7,
    study == "09-0058" & trttrue == "T3" ~ 7,
    study == "10-0017" & trttrue == "T1" ~ 7,
    study == "10-0017" & trttrue == "T2" ~ 7,
    study == "10-0017" & trttrue == "T3" ~ 7,
    study == "13-0033" & trttrue == "T1" ~ 7,
    study == "13-0033" & trttrue == "T2" ~ 7,
    study == "13-0033" & trttrue == "T3" ~ 7,
    study == "15-0064" & trttrue == "1" ~ 7,
    study == "15-0064" & trttrue == "4" ~ 7,
    study == "15-0066" & trttrue == "GROUP1" ~ 7,
    study == "15-0066" & trttrue == "GROUP2" ~ 7,
    study == "17-0075" & trttrue == "GROUP1" ~ 7,
    study == "17-0075" & trttrue == "GROUP2" ~ 7,
    study == "17-0075" & trttrue == "GROUP3" ~ 7,
    study == "17-0078" & trttrue == "1" ~ 7,
    study == "17-0078" & trttrue == "2" ~ 7,
    study == "17-0078" & trttrue == "4" ~ 7,
    study == "17-0078" & trttrue == "5" ~ 7,
    study == "17-0090" & trttrue == "1" ~ 7,
    study == "10-0016" & trttrue == "T1" ~ 8,
    study == "10-0016" & trttrue == "T2" ~ 8,
    study == "10-0016" & trttrue == "T3" ~ 8,
    study == "13-0032" & trttrue == "T1" ~ 8,
    study == "13-0032" & trttrue == "T2" ~ 8,
    study == "13-0032" & trttrue == "T3" ~ 8,
    study == "13-0033" & trttrue == "T8" ~ 8,
    study == "13-0034" ~ 8,
    study == "15-0064" & trttrue == "2" ~ 8,
    study == "15-0064" & trttrue == "5" ~ 8,
    study == "15-0066" & trttrue == "GROUP4" ~ 8,
    study == "15-0066" & trttrue == "GROUP5" ~ 8,
    study == "17-0076" & trttrue == "3.7579WA2x" ~ 8,
    study == "17-0076" & trttrue == "7.579WA2x" ~ 8,
    study == "17-0076" & trttrue == "1579WA2x" ~ 8,
    study == "13-0033" & trttrue == "T4" ~ 3,
    study == "17-0078" & trttrue == "3" ~ 3,
    study == "17-0078" & trttrue == "6" ~ 3,
    study == "13-0033" & trttrue == "T5" ~ 1,
    study == "13-0032" & trttrue == "T4" ~ 4,
    study == "13-0032" & trttrue == "T5" ~ 2,
    study == "13-0033" & trttrue == "T6" ~ 5,
    study == "13-0033" & trttrue == "T7" ~ 6,
    TRUE ~ 0
  ))

# New dose_adj variable
d <- d %>%
  mutate(dosage_adj_2 = case_when(
    dosage == "low" & adj_2 == 0 ~ 0,
    dosage == "low" & adj_2 == 7 ~ 1,
    dosage == "low" & adj_2 == 8 ~ 2,
    dosage == "med" & adj_2 == 0 ~ 3,
    dosage == "med" & adj_2 == 3 ~ 4,
    dosage == "med" & adj_2 == 4 ~ 5,
    dosage == "med" & adj_2 == 6 ~ 6,
    dosage == "med" & adj_2 == 7 ~ 7,
    dosage == "med" & adj_2 == 8 ~ 8,
    dosage == "high" & adj_2 == 0 ~ 9,
    dosage == "high" & adj_2 == 1 ~ 10,
    dosage == "high" & adj_2 == 2 ~ 11,
    dosage == "high" & adj_2 == 3 ~ 12,
    dosage == "high" & adj_2 == 5 ~ 13,
    dosage == "high" & adj_2 == 7 ~ 14,
    dosage == "high" & adj_2 == 8 ~ 15))

cor_stats <- d %>%
  group_by(strain) %>%
  summarise(
    rho = cor(global_react_total, final_logratio, method = "spearman", use = "complete.obs"),
    p = cor.test(global_react_total, final_logratio, method = "spearman")$p.value
  )

cor_stats$p_label <- ifelse(
  cor_stats$p < 0.001,
  "<0.001",
  sprintf("%.3f", cor_stats$p)
)

d$hover_text <- paste0(
  "Reactogenicity: ", d$global_react_total,
  "\nImmunogenicity: ", round(d$final_logratio, 2),
  "\nStrain: ", d$strain
)

p1 <- ggplot(d, aes(x = global_react_total, y = final_logratio)) +
  geom_point(aes(text=hover_text), size = 0.5, alpha = 0.5) +
  facet_wrap(~ strain) +

  geom_text(
    data = cor_stats,
    aes(
      x = 17.5,
      y = -3,
      label = paste0(
        "Rho = ", round(rho, 2),
        "\np = ", p_label
      )
    ),
    inherit.aes = FALSE,
    size = 3
  ) +

  geom_smooth(aes(color = strain), linewidth = 0.5) +
  scale_color_brewer(palette = "Set1") +
  labs(
    x = "Reactogenicity\n(Total number of adverse events)",
    y = "Immunogenicity (Log Ratio)"
  ) +
  theme_minimal() +
  theme(axis.title = element_text(size = 12))+
  theme(legend.position = "none",
        strip.text = element_text(size = 12))

ggplotly(p1, tooltip = "text") %>%
  layout(autosize = TRUE,
           margin = list(
      l = 80,
      r = 30,
      t = 30,
      b = 80
    ))
  
```

### Forest Plot

```{r}
df <- d[d$strain != "H1N1", ]

model <- lmer(
  final_logratio ~ global_react_total + as.factor(age_group) + gender +
  as.factor(dosage_adj_2)*strain +
  (1 + global_react_total | study),
  data = df
)

# Creating data set just for study and strain
study_strain <- d %>%
  select(study, strain) %>%
  distinct()

# Study-specific slopes
re <- ranef(model)$study
re$study <- rownames(re)

beta <- fixef(model)["global_react_total"]

forest_df <- re %>%
  mutate(
    estimate = beta + global_react_total
  ) %>%
  select(study, estimate)

# Standard errors for slopes
re_full <- ranef(model, condVar = TRUE)$study
postvar <- attr(re_full, "postVar")
se <- sqrt(postvar[2,2,])  # slope SE

forest_df$se <- se

forest_df <- forest_df %>%
  mutate(
    lower = estimate - 1.96 * se,
    upper = estimate + 1.96 * se
  )

# Joining with study/strain 
forest_df <- forest_df %>%
  left_join(study_strain, by = "study")

# Pooled estimate
overall <- data.frame(
  study = "Overall",
  estimate = beta,
  se = sqrt(vcov(model)["global_react_total","global_react_total"])
)

overall <- overall %>%
  mutate(
    lower = estimate - 1.96 * se,
    upper = estimate + 1.96 * se
  )

# Combine
forest_df <- bind_rows(forest_df, overall)

# Fix ordering (numeric studies + Overall last)
forest_df <- forest_df %>%
  mutate(
    study = ifelse(study == "Overall", "Overall", as.character(study))
  )

# Order studies by strain, then study
study_order <- forest_df %>%
  filter(study != "Overall") %>%
  arrange(strain, study) %>%
  pull(study)

# Add overall at the end
study_order <- c(study_order, "Overall")

# Apply ordering (reverse for ggplot)
forest_df$study <- factor(forest_df$study, levels = rev(study_order))

# Mark overall vs study
forest_df <- forest_df %>%
  mutate(type = ifelse(study == "Overall", "Overall", "Study"))

# Mark strain for overall
forest_df$strain[forest_df$study == "Overall"] <- "Overall"

# Text for tooltips
forest_df$hover_text <- paste0(
  "Study: ", forest_df$study,
  "<br>Estimate: ", sprintf("%.3f", forest_df$estimate),
  "<br>95% CI: [",
  sprintf("%.3f", forest_df$lower), ", ",
  sprintf("%.3f", forest_df$upper), "]"
)

# Plot
p2 <- ggplot(forest_df,
       aes(x = estimate, y = study)) +
  geom_point(aes(shape = type, size = type, color=strain, text = hover_text)) +
  geom_errorbarh(aes(xmin = lower, xmax = upper, color = strain),
                 height = 0.2) +
  geom_vline(xintercept = 0, color = "#48494B", linewidth = 0.7) +
  geom_vline(xintercept = beta,
             linetype = "dashed",
             color = "black") +
   scale_color_manual(values = c(
    "H5N1" = "#377EB8",
    "H5N8" = "#4DAF4A",
    "H7N9" = "#984EA3",
    "Overall" = "black"
  )) +
  scale_x_continuous(                                 
    breaks = seq(-0.1, 0.15, by = 0.05)
  ) +
  coord_cartesian(xlim = c(-0.1, 0.15)) +
  scale_shape_manual(values = c("Study" = 16, "Overall" = 18)) +  # diamond
  scale_size_manual(values = c("Study" = 3, "Overall" = 4)) +
  labs(
    x = "Study-specific random slopes",
    y = "Study",
    color = "Strain"
  ) +
  theme_bw() +
  theme(axis.title = element_text(size = 12))+
  guides(shape = "none", size = "none")

ggplotly(p2, tooltip="text") %>%
  layout(
    autosize = TRUE,
    margin = list(
      l = 80,
      r = 30,
      t = 30,
      b = 80
    ),
    legend = list(
      title = list(text = "Strain", font = list(size = 11)),
      font=list(size=10),
      itemsizing = "constant",
      x = 0.01,   # left (0 = left edge, 1 = right edge)
      y = 0.99,   # top
      xanchor = "left",
      yanchor = "top",
      bgcolor = "white",
      bordercolor = "black",
      borderwidth = 1
    )
  )
```

Row {data-height=200}
------------------------------------
### Real-World Impact
Understanding the reactogenicity-immunogenicity relationship could improve public health communication by giving meaning to the adverse events that people experience after receiving a vaccination. Additionally, developing reliable predictive frameworks that integrate clinical and immunologic markers could improve early evaluation of vaccine candidates and inform more efficient trial designs.
